---
title: "dplyr"
output: html_notebook
---

```{r setup}
library(dplyr)
library(readr)
library(purrr)
library(fs)
```

# Introduction to `dplyr`

## `dplyr` basics
*Load data into R via `readr`*

Read the **transactions.csv** file, from the **../data** folder. Use the `read_csv()` function to load the data into a variable called `transactions`
```{r}
transactions <- dir_ls("../data", glob = "*.csv") %>%
   map_dfr(read_csv)
```

Preview the data using `glimpse()`
```{r}
glimpse(transactions)
```

## `dplyr` syntax
*How dplyr verbs (functions) can be used to analyze data*

Use `dplyr` verbs to obtain the total sales by month
```{r}
transactions %>%
  group_by(date_month) %>%
  summarise(total_sales = sum(price))
```

Load the above code into a variable called `by_month`
```{r}
by_month <- transactions %>%
  group_by(date_month) %>%
  summarise(total_sales = sum(price))
```
    
Create a new column using `mutate()`
```{r}
transactions %>%
  group_by(order_id) %>% 
  mutate(order_total = sum(price))
```

Count how many observations there are for each date
```{r}
transactions %>% 
  count(date)
```


## Visualization

Load the `ggplot2` library
```{r}
library(ggplot2)
```

Use `as_tibble()` to convert before creating a line plot 
```{r}
by_month %>%
  ggplot() +
  geom_line(aes(date_month, total_sales))
```

## Pivot data
*Review a simple way to aggregate data faster, and then pivot it as a tibble*

Load the `tidyr` library
```{r}
library(tidyr)
```

Group `transactions` by `date_month` and `date_day`, then aggregate `price` into `total_sales`
```{r}
transactions %>%
  group_by(date_month, date_day) %>% 
  summarise(total_sales = sum(price))
```

Copy the aggregation code above and use `pivot_wider()` to make the `date_day` the column headers.
```{r}
transactions %>%
  mutate(date_day = factor(date_day, levels = c(
    "Monday", 
    "Tuesday", 
    "Wednesday", 
    "Thursday", 
    "Friday", 
    "Saturday", 
    "Sunday"
  ))) %>%
  group_by(date_month, date_day) %>% 
  summarise(total_sales = sum(price)) %>%
  pivot_wider(names_from = date_day, values_from = total_sales)
```
