---
title: "arrow"
output: html_notebook
---

```{r setup}
library(arrow)
library(ggplot2)
library(nycflights13)
library(dplyr)
```

# Using `arrow`

Comprehensive examples and documentation for using Apache Arrow with R can be
found at https://arrow.apache.org/docs/r/. This example will specifically explore
working with out-of-memory data in R.

## Connect to the data directory
```{r}
ds <- open_dataset("../data/", format = "csv")
class(ds)
```

## View the `Dataset` object
```{r}
ds
```

## View memory footprint of `ds`
```{r}
lobstr::obj_size(ds)
```

```{r}
local_ds <- collect(ds)
lobstr::obj_size(local_ds)
```

## Use `dplyr` to manipulate the `Dataset` object
Note: *In the current release, arrow supports the dplyr verbs `mutate()`,
`transmute()`, `select()`, `rename()`, `relocate()`, `filter()`, and
`arrange()`. Aggregation is not yet supported, so before you call `summarise()`
or other verbs with aggregate functions, use `collect()` to pull the selected
subset of the data into an in-memory R data frame.*

```{r}
(first_months <- ds %>%
  filter(date_month <= 6))
```

Since summarization functions are not yet supported, this approach is best
suited for filtering and subsetting data to get it to a manageable size prior to
bringing it into R for summarization and further analysis.
