---
title: "Databases"
output: html_notebook
---

```{r setup}
library(DBI)
library(connections)

knitr::opts_chunk$set(connection = "con", max.print = 5)
```

# Introduction to `DBI` 
## Connecting

### `DBI`
Use `dbConnect()` to open a database connection
```{r}
con <- dbConnect(RSQLite::SQLite(), "mydatabase.sqlite")
```

Use `dbDisconnect()` to close a database connection
```{r}
dbDisconnect(con)
```

### `connections`
Use `connection_open()` to open a database connection
```{r}
con <- connection_open(RSQLite::SQLite(), "mydatabase.sqlite")
```

Use `connection_close()` to close a database connection
```{r}
connection_close(con)
```

## Local database basics
*Connecting and adding data to a database*

Use `connection_open()` to open a database connection
```{r}
con <- connection_open(RSQLite::SQLite(), "mydatabase.sqlite")
```
    
Use `dbListTables()` to view existing tables, there should be 0 tables
```{r}
dbListTables(con@con)
```

Use `dbWriteTable()` to create a new table using `mtcars` data. Name it **db_mtcars**
```{r}
dbWriteTable(con, "db_mtcars", mtcars)
```

Use `dbListTables()` to view existing tables, it should return **db_mtcars**
```{r}
dbListTables(con@con)
```

Use `dbGetQuery()` to pass a SQL query to the database
```{r}
dbGetQuery(con, "select * from db_mtcars")
```

Close the database connection using `connection_close()`
```{r}
connection_close(con)
```

## Options for writing tables
*Understand how certain arguments in `dbWriteTable()` work*

Use `connection_open()` to open a Database connection again
```{r}
con <- connection_open(RSQLite::SQLite(), "mydatabase.sqlite")
```

Use `dbWriteTable()` to re-create the **db_mtcars** table using `mtcars` data
```{r, eval = FALSE}
dbWriteTable(con, "db_mtcars", mtcars)
```

```
 Error: Table db_mtcars exists in database, and both overwrite and append are FALSE
```

Use the `append` argument in `dbWriteTable()` to add to the data in the **db_mtcars** table
```{r}
dbWriteTable(con, "db_mtcars", mtcars, append = TRUE)
```

Using `dbGetQuery()`, check the current record count of **db_mtcars** with the following query: "select count() from db_mtcars"
```{r}
dbGetQuery(con, "select count() from db_mtcars")
```

Use the `overwrite` argument to `dbWriteTable()` to replace the data in the **db_mtcars** table
```{r, eval = TRUE}
dbWriteTable(con, "db_mtcars", mtcars, overwrite = TRUE)
```

Check the record count of `db_mtcars` again
```{r}
dbGetQuery(con, "select count() from db_mtcars")
```

## Database operations
*Understand how to use `dbSendStatement()` and `dbExecute()` to modify the database*

Use `dbSendStatement()` to pass a SQL commands that deletes any automatic car from `db_mtcars`: "delete from db_mtcars where am = 1". Load the command to a variable called `rs`
```{r}
rs <- dbSendStatement(con, "delete from db_mtcars where am = 1")
```

Call the `rs` variable to view information about the results of the requested change
```{r}
rs
```

Use `dbHasCompleted()` to confirm that the job is complete
```{r}
dbHasCompleted(rs)
```

Use `dbGetRowsAffected()` to see the number of rows that were affected by the request
```{r}
dbGetRowsAffected(rs)
```

Clear the results using `dbClearResult()`
```{r}
dbClearResult(rs)
```

Confirm that the result set has been removed by calling the `rs` variable once more
```{r}
rs
```

Check the record count of **db_mtcars** again, the new count should be 19 (32 original records - 13 deleted records)
```{r}
dbGetQuery(con, "select count() from db_mtcars")
```

Use `dbWriteTable()` to overwrite **db_mtcars** with the value of `mtcars`
```{r}
dbWriteTable(con, "db_mtcars", mtcars, overwrite = TRUE)
```
    
Use `dbExeceute()` to delete rows where am = 1 using the same query as before. Load the results in a variable called `rs`
```{r}
rs <- dbExecute(con, "delete from db_mtcars where am = 1")
```
    
`rs` contains the number of rows affected by the statement that was executed
```{r}
rs
```
    
Check the record count of **db_mtcars** again.
```{r}
dbGetQuery(con, "select count() from db_mtcars")
```

## `knitr` SQL engine
*See how to run SQL queries as code chunks*

Start a new code chunk, but using `sql` instead of `r` as the first argument of the chunk.  Also add `connection=con` as another argument of the chunk.

    `r ''````{sql, connection=con}
    select * from db_mtcars
    `r ''````

```{sql, connection=con, echo = FALSE}
select * from db_mtcars
```

Add the `max.print` options to the chunk, and set it to 5

    `r ''````{sql, connection=con, max.print = 5}
    select * from db_mtcars
    `r ''````

```{sql, connection=con, echo = FALSE, max.print = 5}
select * from db_mtcars
```

Set defaults for the `sql` chunks by using the `knitr::opts_chunk$set()` command in the `setup` at the beginning of the document.
    `r ''````{r setup}
    knitr::opts_chunk$set(connection = "con", max.print = 5)
    `r ''````

Run the same query in a new `sql` chunk, but without any other argument

    `r ''````{sql}
    select * from db_mtcars
    `r ''````

```{sql, echo = FALSE}
select * from db_mtcars
```

Store the results of the query into an R object called `local_mtcars` using 
the `output.var` option.
```{sql, output.var = "local_mtcars"}
select * from db_mtcars
```

```{r}
local_mtcars
```

Close the database connection using `connection_close()`
```{r, eval = TRUE}
connection_close(con)
```