---
title: "dbplyr"
output: html_notebook
---

```{r setup}
library(connections)
library(RSQLite)
library(dplyr)
library(dbplyr)
library(config)
```

# Databases and `dplyr`

## Intro to `connections`
*Use `connections` to open open a database connection*

Use `connection_open()` to open a Database connection
```{r}
con <- connection_open(
  RPostgres::Postgres(),
  host =  get("host"),
  user = get("user"),
  password = get("pwd"),
  port = get("port"),
  dbname = "retail",
  bigint = "integer"
)
```
    
The RStudio Connections pane should show the tables in the database.

## Table reference
*Use the `dplyr`'s `tbl()` command*

Load the `dplyr` package
```{r}
library(dplyr)
```

Add `in_schema()` as an argument to `tbl()` to specify the schema
```{r}
tbl(con, "customer")
```

Load the results from the `tbl()` command that points the table called **orders** to a variable called `orders` 
```{r}
orders <- tbl(con, "orders")
```
    
Use the `class` function to determine the object type of `orders`
```{r}
class(orders)
```

## Under the hood 
*Use `show_query()` to preview the SQL statement that will be sent to the database*

Use `show_query()` to preview SQL statement that actually runs when we run `orders` as a command
```{r}
show_query(orders)
```
    
When executed, `orders` returns the first 1000 rows of the remote **orders** table
```{r}
orders
```

Full results of a remote query can be brought into R with `collect`
```{r}
local_orders <- collect(orders)
```

Easily view the resulting query by adding `show_query()` in another piped command
```{r}
orders %>%
  show_query()
```

Insert `head()` in between the two statements to see how the SQL changes
```{r}
orders %>%
  head() %>%
  show_query()
```
    
Queries can be assigned to variables. Create a variable called `orders_head` that contains the previous query
```{r}
orders_head <- orders %>% 
  head()

orders_head %>% 
  show_query()
```

Use `sql_render()` and `simulate_mssql()` to see how the SQL statement changes from vendor to vendor
```{r}
orders %>%
  head() %>%
  sql_render(con = simulate_mssql()) 
```

Use `explain()` to explore the query plan
```{r}
orders %>% 
  head() %>% 
  explain()
```
    
## Un-translated R commands
*Review of how `dbplyr` handles R commands that have not been translated into a like-SQL command*

Preview how `mean` is translated
```{r}
orders %>% 
  mutate(avg_id = mean(order_id, na.rm = TRUE)) %>% 
  show_query()
```

Preview how `Sys.Date()` is translated
```{r}
orders %>%
  mutate(today = Sys.Date()) %>%
  show_query()
```

Use PostgreSQL native commands, in this case `date`
```{r}
orders %>%
  mutate(today = date('now')) %>%
  show_query()
```

Run the `dplyr` code to confirm it works
```{r}
orders %>%
  mutate(today = date('now')) %>%
  head()
```

## Using bang-bang
*Intro on passing unevaluated code to a dplyr verb*

Preview how `Sys.Date()` is translated when prefixing `!!`
```{r}
orders %>%
  mutate(today = !!Sys.Date()) %>%
  show_query()
```

View resulting table when `Sys.Date()` is translated when prefixing `!!`
```{r}
orders %>%
  mutate(today = !!Sys.Date()) %>%
  head()
```

Disconnect from the database using `connection_close`
```{r}
connection_close(con)
```

