---
title: "dtplyr"
output: html_notebook
---

```{r setup}
library(data.table)
library(dtplyr)
library(dplyr)
library(lobstr)
library(fs)
library(purrr)
```

# Introduction to `dtplyr`

## `dtplyr` basics
*Load data into R via `data.table`, and then wrap it with `dtplyr`*

Read the **transactions.csv** file, from the **../data** folder. Use the `fread()` function to load the data into a variable called `transactions`
```{r}
transactions <- dir_ls("../data", glob = "*.csv") %>%
   map(fread) %>%
   rbindlist()
```

Preview the data using `glimpse()`
```{r}
glimpse(transactions)
```

Use `lazy_dt()` to "wrap" the `transactions` variable into a new variable called `dt_transactions`
```{r}
dt_transactions <- lazy_dt(transactions)
```

View `dt_transactions` structure with `glimpse()`
```{r}
glimpse(dt_transactions)
```

## Object sizes
*Confirm that `dtplyr` is not making copies of the original `data.table`*

Load the `lobstr` library
```{r}
library(lobstr)
```

Use `obj_size()` to obtain `transactions`'s size in memory
```{r}
obj_size(transactions)
```

Use `obj_size()` to obtain `dt_transactions`'s size in memory
```{r}
obj_size(dt_transactions)
```

Use `obj_size()` to obtain `dt_transactions` and `transactions` size in memory together
```{r}
obj_size(transactions, dt_transactions)
```

## How `dtplyr` works
*Under the hood view of how `dtplyr` operates `data.table` objects*

Use `dplyr` verbs on top of `dt_transactions` to obtain the total sales by month
```{r}
dt_transactions %>%
  group_by(date_month) %>%
  summarise(total_sales = sum(price))
```

Load the above code into a variable called `by_month`
```{r}
by_month <- dt_transactions %>%
  group_by(date_month) %>%
  summarise(total_sales = sum(price))
```

Use `show_query()` to see the `data.table` code that `by_month` actually runs
```{r}
show_query(by_month)
```

Use `glimpse()` to view how `by_month`, instead of modifying the data, only adds steps that will later be executed by `data.table`
```{r}
glimpse(by_month)
```
    
Create a new column using `mutate()`
```{r}
dt_transactions %>%
  mutate(new_field = price / 2)
```

Use `show_query()` to see the `copy()` command being used
```{r}
dt_transactions %>%
  mutate(new_field = price / 2) %>%
  show_query()
```

Check to confirm that the new column *did not* persist in `dt_transactions`
```{r}
dt_transactions
```

Use `lazy_dt()` with the `immutable` argument set to `FALSE` to avoid the copy
```{r}
m_transactions <- lazy_dt(copy(transactions), immutable = FALSE)
```

```{r}
m_transactions
```

Create a `new_field` column in `m_transactions` using `mutate()`
```{r}
m_transactions %>% 
  mutate(new_field = price / 2)
```

Use `show_query()` to see that `copy()` is no longer being used
```{r}
m_transactions %>% 
  mutate(new_field = price / 2) %>% 
  show_query()
```

Inspect `m_transactions` to see that `new_field` has persisted
```{r}
m_transactions
```

## Working with `dtplyr`
*Learn data conversion and basic visualization techniques*

Use `as_tibble()` to convert the results of `by_month` into a `tibble`
```{r}
by_month %>%
  as_tibble()
```

Load the `ggplot2` library
```{r}
library(ggplot2)
```

Use `as_tibble()` to convert before creating a line plot 
```{r}
by_month %>%
  as_tibble() %>%
  ggplot() +
  geom_line(aes(date_month, total_sales))
```

## Pivot data
*Review a simple way to aggregate data faster, and then pivot it as a tibble*

Load the `tidyr` library
```{r}
library(tidyr)
```

Group `db_transactions` by `date_month` and `date_day`, then aggregate `price` into `total_sales`
```{r}
dt_transactions %>%
  group_by(date_month, date_day) %>% 
  summarise(total_sales = sum(price))
```

Copy the aggregation code above, **collect it into a `tibble`**, and then use `pivot_wider()` to make the `date_day` the column headers.
```{r}
dt_transactions %>%
  mutate(date_day = factor(date_day, levels = c(
    "Monday", 
    "Tuesday", 
    "Wednesday", 
    "Thursday", 
    "Friday", 
    "Saturday", 
    "Sunday"
  ))) %>%
  group_by(date_month, date_day) %>% 
  summarise(total_sales = sum(price)) %>%
  as_tibble() %>%
  pivot_wider(names_from = date_day, values_from = total_sales)
```
